Eco-evolutionary dynamics of finite populations from first principles

### 2023-04-24 ind_files/06_nontechnical_discussion.tex
Minor rephrasing

~\cite{lion_theoretical_2018} has recently proposed a reformulation of the relatively well-known unification of eco-evolutionary dynamics via the Price equation~\citep{frank_natural_2012, queller_fundamental_2017, luque_mirror_2021} in a dynamically sufficient, continuous time framework using `feedback' as a `unifying perspective'. Our equations generalize the unifying framework described in~\cite{lion_theoretical_2018} to finite, fluctuating populations --- taking $K \to \infty$ in equations \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price}, and \eqref{nD_stochastic_Price_variance} recover equations (6), (11), and (14) in~\cite{lion_theoretical_2018} respectively. \cite{lion_theoretical_2018} has pointed out that in the dynamic setting (for infinite populations), the replicator-mutator equation \eqref{nD_replicator_mutator} is in some sense the `most fundamental' of the lot, and equations like the Price equation are best viewed as a hierarchy of moment equations for the population mean, population variance, etc. of a type-level quantity. This is also true in our framework - equation \eqref{nD_stochastic_RM} is the fundamental equation for population dynamics, and equations like \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} can then be derived from \eqref{nD_stochastic_RM} through repeated application of It\^o's formula (in principle for any moment of the distribution of $f$ in the population, though this quickly becomes too tedious to actually carry out in practice). If we assume that the quantity $f$ follows a Gaussian distribution, then the mean and variance completely characterize the distribution,  and equations \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price}, and \eqref{nD_stochastic_Price_variance} together specify the complete stochastic dynamics of the system.

>>>>

~\cite{lion_theoretical_2018} has recently proposed a reformulation of the relatively well-known unification of eco-evolutionary dynamics via the Price equation~\citep{frank_natural_2012, queller_fundamental_2017, luque_mirror_2021} in a dynamically sufficient, continuous time framework using `feedback' as a `unifying perspective'. Our equations generalize the unifying framework described in~\cite{lion_theoretical_2018} to finite, fluctuating populations --- taking $K \to \infty$ in equations \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price}, and \eqref{nD_stochastic_Price_variance} recover equations (6), (11), and (14) in~\cite{lion_theoretical_2018} respectively. \cite{lion_theoretical_2018} has pointed out that in the dynamic setting (for infinite populations), the replicator-mutator equation \eqref{nD_replicator_mutator} is in some sense the truly `fundamental' equation for evolutionary dynamics, and equations like the Price equation are best viewed as a hierarchy of moment equations for the population mean, population variance, etc. of a type-level quantity. This is also true in our framework - equation \eqref{nD_stochastic_RM} is the fundamental equation for population dynamics, and equations like \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} can then be derived from \eqref{nD_stochastic_RM} through repeated application of It\^o's formula (in principle for any moment of the distribution of $f$ in the population, though this quickly becomes too tedious to actually carry out in practice). If we assume that the quantity $f$ follows a Gaussian distribution, then the mean and variance completely characterize the distribution,  and equations \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price}, and \eqref{nD_stochastic_Price_variance} together specify the complete stochastic dynamics of the system.
